Light inextensible strings under tension in the Schwarzschild geometry

### 2019-02-07 string_ajp.tex
more editing for AJP

This would indicate that, close to the Schwarzschild radius $r=1$,
large (but finite) amounts of string are required to connect two
points with slightly differing radial coordinates.  Passing to the
case $w\gg 1$ we have

@@

\noindent suggesting that there is ``extra radial space'', of length
over $1.19$ Schwarzschild radii, sensible at large distances from the
event horizon.  This might suggest that radial strings have to be
longer than expected.

@@

This differential equation does not appear to be soluble analytically;
solutions correspond to a taut string confined to Flamm's
paraboloid~\cite{flamm1916}; see figure~\ref{flamm}.  It is
straightforward to show that the string has a point of inflection
if~$\left(r'\right)^2=2r\left(r-1\right)$.

@@

@@

\begin{figure}[h!] % Created by "closest_approach.R"
\centering
\includegraphics[width=70mm]{closest_approach_nonselfintersecting.pdf}
\caption{Non-self intersecting strings arranged by distance~$d$ of
  closest approach to the event horizon, occurring tangentially
  at~$\phi=0$}
\label{closest_approach_non_self_intersecting}
\end{figure}

\begin{figure}[h!] % created by "closest_approach2.R"
\centering
\includegraphics[width=70mm]{closest_approach_selfintersecting.pdf}
\caption{Strings arranged by closest approach to the event horizon:
 self-intersecting strings shown for $\phi\geq 0$}
\label{closest_approach_self_intersecting}
\end{figure}

\begin{figure}[h!] % created by "one_free_end_fixed_EH_intersection.R"
\centering
\includegraphics[width=70mm]{one_free_end_fixed_EH_intersection.pdf}
\caption{Strings tangential to event horizon}
\label{fixed_EH_intersection}
\end{figure}

>>>>

This would indicate that, close to the Schwarzschild radius $r=1$,
large (but finite) amounts of string are required to connect two
points with slightly differing radial coordinates.  Passing to the
case $w\gg 1$ we have an asymptotic expansion that begins

@@

\noindent suggesting that there is ``extra radial space'', of length
over $1.19$ Schwarzschild radii, sensible at large distances from the
event horizon.  This might suggest that radial strings have longer
proper length than expected.

@@

@@

Solutions to this differential equation correspond to a taut string
confined to Flamm's paraboloid~\cite{flamm1916}; see
figure~\ref{flamm}.

There does not appear to be a simple analytical solution but
nevertheless we may make several observations.
Equation~\ref{rdashdash} is a second-order nonlinear ordinary
differential equation and thus has two constants of integration.
There are two qualitatively different types of solutions: those with a
tangential point ($r'=0,r>1$), and those without.  Consider first
solutions with such a point: there exists a pair~$\left(r,\phi\right),
r>1$ at which~$r'(\phi)=0$.  By Rolle's theorem, this point must be
unique, as~$r''(\phi)>0$, and be the closest approach of the string to
the event horizon.

Further, because the system is symmetrical
under~$\phi\longrightarrow-\phi$, the string must be symmetrical about
at this point.  The (signed) radius of curvature~$R$ is readily
evaluated:

\begin{equation}\label{roc}
  R = \frac{
    \left(r^2 + \left(r'\right)^2\right)^\frac{3}{2}
  }{
%    \left|
    r^2 + 2\left(r'\right)^2-rr''
%    \right|
  }
  =
  \frac{
    \left(r^2 + \left(r'\right)^2\right)^\frac{3}{2}
  }{
    r-\frac{\left(r'\right)^2}{2\left(r-1\right)}
  }\qquad\mbox{along a taut string}
\end{equation}

\noindent which would give~$R$ as a function of~$\left(r,r'\right)$.
Considering~$r'(\phi)=0$ shows that if a string is tangential at a
point with~$r_0>1$, then it has a radius of curvature of~$r_0^2$ at
that point.  Any circular path, including a circular orbit at~$r=r_0$,
has a radius of curvature of $r_0<r_0^2$, showing that circular orbits
cannot follow a taut string at closest approach.  Also observe that
null geodesics have a radius of curvature of $2r_0^2/3 < r_0^2$,
showing that light cannot follow a taut string at closest approach
either.

Compare the above analysis with a taut {\em heavy} string which
adopts, at least locally, a catenary configuration at closest
approach~\cite{nguyen2007}.  Such a configuration is manifestly convex
to the singularity, in contrast to light strings which are concave.

If, conversely, $r'(\phi)>0$ whenever~$r>1$, then the string has only
one free end; the other is tangential to the event horizon, except in
the exceptional case of constant~$\phi$ when the string is radial
for~$r>1$.

@@

\begin{figure}[h!] % Created by "closest_approach.R"
\centering
\includegraphics[width=170mm]{closest_approach_nonselfintersecting.pdf}
\caption{Non-self intersecting strings arranged by distance~$d$ of
  closest approach to the event horizon, occurring tangentially
  at~$\phi=0$}
\label{closest_approach_non_self_intersecting}
\end{figure}

\begin{figure}[h!] % created by "closest_approach2.R"
\centering
\includegraphics[width=170mm]{closest_approach_selfintersecting.pdf}
\caption{Strings arranged by closest approach to the event horizon:
 self-intersecting strings shown for $\phi\geq 0$}
\label{closest_approach_self_intersecting}
\end{figure}

\begin{figure}[h!] % created by "one_free_end_fixed_EH_intersection.R"
\centering
\includegraphics[width=170mm]{one_free_end_fixed_EH_intersection.pdf}
\caption{Strings tangential to event horizon}
\label{fixed_EH_intersection}
\end{figure}
